---
title: 'Lab Report: Bike-Sharing Data Analysis'
subtitle: "MA 575 Fall 2021 - C3 Team #2"
author: "Ali Taqi, Hsin-Chang Lin, Huiru Yang, Ryan Mahoney, Yulin Li"
date: "12/01/2021"
output: 
  pdf_document: 
    number_sections: true
---

## Preprocessing

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
#==============================#
#       Import Libraries       #
#==============================#
library(tidyverse)
library(GGally)
library(cowplot)
library(gglm)
#==============================#
#        Global Settings       #
#==============================#
# Center ggplot titles
theme_update(plot.title = element_text(hjust = 0.5))
#=============================#
#        Obtain Dataset       #
#=============================#
# Source the wrangling scripts
source(file = "../R/wrangle-lynne.R")
# Import and wrangle data
bike_day <- read.csv("../data/bike-day.csv", header = T)
# Use the wrangling "wrapper" function to clean the data (fxn sourced from wrangle.R)
data <- wrangle_init(bike_day)
# Partition the 2011/2012 data
data2011 <- in_2011(data); data2012 <- in_2012(data)
#============================#
#       Source Scripts       #
#============================#
# Import any models (and fitted values, if applicable)
source(file = "../R/model-lynne.R")
# Import any plots or plot wrapper functions
source(file = "../R/plot-lynne.R")
```

## Visualization

### Pairwise Relationships

$$ $$

**Group 1: (Mainly) Environmental**

Predictors (all environmental variables plus date):

- `dteday` (date)
- `weathersit` (weather type)
- `temp` (measured temperature)
- `atemp` (feeling temperature)
- `windspeed` (wind speed)
- `hum` (humidity)

Response:

- `cnt` (daily total count)
- `registered` (daily registered count)
- `casual` (daily casual count)

$$ $$

**Group 2: Seasonal**

Predictors (all seasonal variables):

- `dteday` (date)
- `season` (season)
- `yr` (year)
- `mnth` (month)
- `holiday` (holiday or not)
- `weekday` (day of the week)
- `workingday` (working day or not)

Response:

- `cnt` (daily total count)
- `registered` (daily registered count)
- `casual` (daily casual count)

$$ $$

#### Pairs Plot for Group 1 (Mainly Environmental)

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs(data, group = 1, y_name = 'cnt')
plot_pairs(data, group = 1, y_name = 'registered')
plot_pairs(data, group = 1, y_name = 'casual')
```

**Observations:**
 - `registered`: quadratic in `temp`
 - `casual`: linear in `temp`
 

#### Pairs Plot for Group 1 (Mainly Environmental) (Colored by Year)

```{r}
# plot_pairs_colored(data, group = 0, y_name = 'cnt', color_name = 'yr', var_list_new = c('mnth', 'atemp'))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs_colored(data, group = 1, y_name = 'cnt', color_name = 'yr')
plot_pairs_colored(data, group = 1, y_name = 'registered', color_name = 'yr')
plot_pairs_colored(data, group = 1, y_name = 'casual', color_name = 'yr')
```

**Observations:**
 - `registered`: sensitive to `yr`
 - `casual`: less sensitive to `yr`

#### Pairs Plot for Group 1 (Mainly Environmental) (Colored by Weather Type)

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs_colored(data, group = 1, y_name = 'cnt', color_name = 'weathersit')
plot_pairs_colored(data, group = 1, y_name = 'registered', color_name = 'weathersit')
plot_pairs_colored(data, group = 1, y_name = 'casual', color_name = 'weathersit')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs_colored(data2011, group = 1, y_name = 'cnt', color_name = 'weathersit')
plot_pairs_colored(data2011, group = 1, y_name = 'registered', color_name = 'weathersit')
plot_pairs_colored(data2011, group = 1, y_name = 'casual', color_name = 'weathersit')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs_colored(data2012, group = 1, y_name = 'cnt', color_name = 'weathersit')
plot_pairs_colored(data2012, group = 1, y_name = 'registered', color_name = 'weathersit')
plot_pairs_colored(data2012, group = 1, y_name = 'casual', color_name = 'weathersit')
```

#### Pairs Plot for Group 2 (Seasonal)

```{r}
# plot_pairs_colored(data, group = 0, color_name = 'yr', y_name = 'cnt', var_list_new = c('mnth'))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs(data, group = 2, y_name = 'cnt')
plot_pairs(data, group = 2, y_name = 'registered')
plot_pairs(data, group = 2, y_name = 'casual')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_pairs_colored(data, group = 2, color_name = 'yr', y_name = 'cnt')
plot_pairs_colored(data, group = 2, color_name = 'yr', y_name = 'registered')
plot_pairs_colored(data, group = 2, color_name = 'yr', y_name = 'casual')
```

**Observations:**

 - `casual`: `workingday` seems significant
 - `registered`: `workingday` seems less significant
