---
title: 'Lab Report: Bike-Sharing Data Analysis'
subtitle: "MA 575 Fall 2021 - C3 Team #2"
author: "Ali Taqi, Hsin-Chang Lin, Huiru Yang, Ryan Mahoney, Yulin Li"
date: "12/01/2021"
output: 
  pdf_document: 
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.height = 4, fig.width = 6)
#==============================#
#       Import Libraries       #
#==============================#
library(tidyverse)
library(GGally)
library(cowplot)
library(gglm)
library(moderndive)
#==============================#
#        Global Settings       #
#==============================#
# Center ggplot titles
theme_update(plot.title = element_text(hjust = 0.5))
#=============================#
#        Obtain Dataset       #
#=============================#
# Source the wrangling scripts
source(file = "../R/wrangle-lynne.R")
# Import and wrangle data
bike_day <- read.csv("../data/bike-day.csv", header = T)
# Use the wrangling "wrapper" function to clean the data (fxn sourced from wrangle.R)
data <- wrangle_init(bike_day)
# Partition the 2011/2012 data
data2011 <- in_2011(data); data2012 <- in_2012(data)
#============================#
#       Source Scripts       #
#============================#
# Import any models (and fitted values, if applicable)
source(file = "../R/model-lynne.R")
# Import any plots or plot wrapper functions
source(file = "../R/plot.R")
source(file = "../R/plot-lynne.R")
```

# Data Visualization

## Plots

```{r}
cols <- c('all day types' = "darkgrey", 
          'on non-working days' = "darkgreen", 
          'on holidays' = "red", 
          'trends' = 'black', 
          'fitted' = 'purple')
```

\newpage

### 2011 Data

```{r fig.width=6, fig.height=4, echo = FALSE}
# Plot 1: 2011, Registered
#plot_usertype_cnt(data2011, user = "registered")
p1 <- plot_usertype_cnt(data2011, user = "registered", compare = T)
p1
# Plot 2: 2011, Casual
#plot_usertype_cnt(data2011, user = "casual")
p2 <- plot_usertype_cnt(data2011, user = "casual", compare = T)
p2
```

\newpage

### 2012 Data

```{r fig.width=6, fig.height=4, echo = FALSE}
# Plot 3: 2012, Causal
#plot_usertype_cnt(data2012, user = "casual")
p3 <- plot_usertype_cnt(data2012, user = "casual", compare = T)
p3
# Plot 4: 2012, Registered
#plot_usertype_cnt(data2012, user = "registered")
p4 <- plot_usertype_cnt(data2012, user = "registered", compare = T)
p4
```


- casual counts & registered counts: opposite behaviors - active on weekends vs active on working days. 


\newpage

# Modeling

- for casual counts: adding temp^2 makes the prediction even more unstable.
- for registered counts: adding temp^2 seems necessary. 

```{r}
### Unused models:
mod.cas.1 <- lm(casual ~
                   workingday + weathersit + atemp + I(atemp^2), data = data2011)
mod.reg.1 <- lm(registered ~
                   workingday + weathersit + atemp + I(atemp^2), data = data2011)
mod.cas.2 <- lm(casual ~
                   weathersit:workingday:atemp + season, data = data2011)
mod.reg.2 <- lm(registered ~
                   weathersit:workingday:atemp + I(atemp^2) + weathersit + season, data = data2011)
mod.cas.3 <- lm(casual ~
                   workingday:atemp + weathersit:season, data = data2011)
mod.reg.3 <- lm(registered ~
                   workingday:atemp + I(atemp^2) + weathersit + season, data = data2011)
mod.reg.4 <- lm(registered ~
                  season:workingday:atemp + season:workingday:I(atemp^2) + weathersit, data = data2011)
# summary(mod.cas.1)
# summary(mod.reg.1)
# summary(mod.cas.2)
# summary(mod.reg.2)
# summary(mod.reg.4)
```

## Casual Users

```{r, echo = T}
mod.cas.3 <- lm(casual ~ season:workingday:atemp + weathersit, data = data2011)
```

```{r}
# Current best performing model: mod.cas.3
get_regression_table(mod.cas.3)
get_regression_summaries(mod.cas.3) 
```

\newpage

## Registered Users

```{r, echo = T}
mod.reg.3 <- lm(registered ~ season:workingday:atemp + season:I(atemp^2) + weathersit, data = data2011)
```

```{r}
get_regression_table(mod.reg.3)
get_regression_summaries(mod.reg.3)
```

\newpage

# Diagnostics: 2011 Data

## Our Wrapper Functions

```{r, echo = T}
plot_fitted_cnt_2011 <- function(data, user, model, compare = FALSE){
  plot_usertype_cnt(data, user, compare) + 
    geom_line(mapping = aes(x = dteday, y = predict(model, data2011), color = 'fitted'), 
              size = 0.5, alpha = 0.7)
}
```

\newpage

## Casual Users

### Model 2

```{r}
plot_fitted_cnt_2011(data = data2011, user = "casual", model = mod.cas.2, compare = TRUE)
```


### Model 3

```{r, echo = T}
plot_fitted_cnt_2011(data = data2011, user = "casual", model = mod.cas.3, compare = TRUE)
```

\newpage

## Registered Users

### Model 2

```{r}
plot_fitted_cnt_2011(data = data2011, user = "registered", model = mod.reg.2, compare = TRUE)
```

### Model 3

```{r, echo = T}
plot_fitted_cnt_2011(data = data2011, user = "casual", model = mod.cas.3, compare = TRUE)
```

### Model 4

```{r, echo = T}
#plot_fitted_cnt_2011(data = data2011, user = "casual", model = mod.cas.4, compare = TRUE)
```

\newpage

# Diagnostics: 2012 Data

## Our Wrapper Functions

```{r, echo = T}
plot_fitted_cnt_2012 <- function(data, user, model, compare = FALSE){
  plot_usertype_cnt(data, user, compare) + 
    geom_line(data = data2012, mapping = aes(x = dteday, y = predict(model, data2012), color = 'fitted'), 
              size = 0.5, alpha = 0.7)
}
```

\newpage

## Registered Users

### Model 2

```{r}
plot_fitted_cnt_2012(data = data2012, user = "registered", model = mod.reg.2, compare = FALSE)
```

### Model 3

```{r}

```

\newpage

## Casual Users

### Model 2

```{r}

```

### Model 3

```{r}

```

```{r}
fitted.reg <- predict(mod.reg.3, data2012)
fitted.cas <- predict(mod.cas.3, data2012)

p4 + geom_line(data2012, 
            mapping = aes(dteday, fitted.cas/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0) + 
  geom_line(data2012, 
            mapping = aes(dteday, fitted.cas/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.8)

p3 + geom_line(data2012, 
            mapping = aes(dteday, predict(mod.reg.4, data2012)/0.608, color = 'on holidays'), 
            size = 0.5, alpha = 0) + 
  geom_line(data2012, 
            mapping = aes(dteday, predict(mod.reg.3, data2012)/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.8)
```

#### Model 2: current best

```{r}
fitted.reg <- predict(mod.reg.2, data2012)
fitted.cas <- predict(mod.cas.2, data2012)

p4 + geom_line(data2012, 
            mapping = aes(dteday, fitted.cas/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.7)

p3 + geom_line(data2012, 
            mapping = aes(dteday, fitted.reg/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.7)
```

