---
title: 'Lab Report: Bike-Sharing Data Analysis'
subtitle: "MA 575 Fall 2021 - C3 Team #2"
author: "Ali Taqi, Hsin-Chang Lin, Huiru Yang, Ryan Mahoney, Yulin Li"
date: "12/01/2021"
output: 
  pdf_document: 
    number_sections: true
---

## Preprocessing

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
#==============================#
#       Import Libraries       #
#==============================#
library(tidyverse)
library(GGally)
library(cowplot)
library(gglm)
#==============================#
#        Global Settings       #
#==============================#
# Center ggplot titles
theme_update(plot.title = element_text(hjust = 0.5))
#=============================#
#        Obtain Dataset       #
#=============================#
# Source the wrangling scripts
source(file = "../R/wrangle-lynne.R")
# Import and wrangle data
bike_day <- read.csv("../data/bike-day.csv", header = T)
# Use the wrangling "wrapper" function to clean the data (fxn sourced from wrangle.R)
data <- wrangle_init(bike_day)
# Partition the 2011/2012 data
data2011 <- in_2011(data); data2012 <- in_2012(data)
#============================#
#       Source Scripts       #
#============================#
# Import any models (and fitted values, if applicable)
source(file = "../R/model-lynne.R")
# Import any plots or plot wrapper functions
source(file = "../R/plot-lynne.R")
```

## Plots

```{r}
cols <- c('all day types' = "darkgrey", 
          'on non-working days' = "darkgreen", 
          'on holidays' = "red", 
          'trends' = 'black', 
          'fitted' = 'purple')
```


```{r fig.width=7, fig.height=4, echo=FALSE}
### report 4

p1 <- 
  ggplot(data2011, aes(dteday, registered, color = 'all day types')) + 
  geom_point(size = 1.5, pch = 20, alpha = 0.8) + 
  labs(x = 'Date', y = 'Daily Registered Count') + 
  
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5)) +
  
  geom_line(data2011, 
            mapping = aes(dteday, registered, color = 'trends'), 
            size = 0.3, alpha = 0.8) + 

  geom_point(data2011[!(data2011$workingday), ], 
             mapping = aes(dteday, registered, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 1) + 
  geom_point(data2011[data2011$holiday, ], 
             mapping = aes(dteday, registered, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 1) + 
  scale_color_manual(values = cols, 
                     labels = names(cols), 
                     name = 'Daily Count')

p1 
```

```{r}
p1 <- 
  p1 + geom_line(data2011, 
            mapping = aes(dteday, casual, color = 'trends'), 
            size = 0.3, alpha = 0.2) + 

  geom_point(data2011[!(data2011$workingday), ], 
             mapping = aes(dteday, casual, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 0.2) + 
  geom_point(data2011[data2011$holiday, ], 
             mapping = aes(dteday, casual, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 0.2)

p1
```


```{r fig.width=7, fig.height=4, echo=FALSE}
### report 4

p2 <-
  ggplot(data2011, aes(dteday, casual, color = 'all day types')) + 
  geom_point(size = 1.5, pch = 20, alpha = 0.8) + 
  labs(x = 'Date', y = 'Daily Registered Count') + 
  
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5)) +
  
  geom_line(data2011, 
            mapping = aes(dteday, casual, color = 'trends'), 
            size = 0.3, alpha = 0.8) + 

  geom_point(data2011[!(data2011$workingday), ], 
             mapping = aes(dteday, casual, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 1) + 
  geom_point(data2011[data2011$holiday, ], 
             mapping = aes(dteday, casual, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 1) + 
  scale_color_manual(values = cols, 
                     labels = names(cols), 
                     name = 'Daily Count')

p2 <- p2 + 
  geom_line(data2011, 
            mapping = aes(dteday, registered, color = 'trends'), 
            size = 0.3, alpha = 0.2) + 

  geom_point(data2011[!(data2011$workingday), ], 
             mapping = aes(dteday, registered, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 0.2) + 
  geom_point(data2011[data2011$holiday, ], 
             mapping = aes(dteday, registered, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 0.2)

p2
```

```{r}
p3 <- 
  ggplot(data2012, aes(dteday, registered, color = 'all day types')) + 
  geom_point(size = 1.5, pch = 20, alpha = 0.8) + 
  labs(x = 'Date', y = 'Daily Registered Count') + 
  
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5)) +
  
  geom_line(data2012, 
            mapping = aes(dteday, registered, color = 'trends'), 
            size = 0.3, alpha = 0.8) + 

  geom_point(data2012[!(data2012$workingday), ], 
             mapping = aes(dteday, registered, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 1) + 
  geom_point(data2012[data2012$holiday, ], 
             mapping = aes(dteday, registered, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 1) + 
  scale_color_manual(values = cols, 
                     labels = names(cols), 
                     name = 'Daily Count')
p3 <- 
  p3 + geom_line(data2012, 
            mapping = aes(dteday, casual, color = 'trends'), 
            size = 0.3, alpha = 0.2) + 

  geom_point(data2012[!(data2012$workingday), ], 
             mapping = aes(dteday, casual, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 0.2) + 
  geom_point(data2012[data2012$holiday, ], 
             mapping = aes(dteday, casual, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 0.2)

p4 <- 
  ggplot(data2012, aes(dteday, casual, color = 'all day types')) + 
  geom_point(size = 1.5, pch = 20, alpha = 0.8) + 
  labs(x = 'Date', y = 'Daily Registered Count') + 
  
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5)) +
  
  geom_line(data2012, 
            mapping = aes(dteday, casual, color = 'trends'), 
            size = 0.3, alpha = 0.8) + 

  geom_point(data2012[!(data2012$workingday), ], 
             mapping = aes(dteday, casual, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 1) + 
  geom_point(data2012[data2012$holiday, ], 
             mapping = aes(dteday, casual, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 1) + 
  scale_color_manual(values = cols, 
                     labels = names(cols), 
                     name = 'Daily Count')

p4 <- p4 + 
  geom_line(data2012, 
            mapping = aes(dteday, registered, color = 'trends'), 
            size = 0.3, alpha = 0.2) + 

  geom_point(data2012[!(data2012$workingday), ], 
             mapping = aes(dteday, registered, color = 'on non-working days'), 
             size = 1.5, pch = 16, alpha = 0.2) + 
  geom_point(data2012[data2012$holiday, ], 
             mapping = aes(dteday, registered, color='on holidays'), 
             size = 1.5, pch = 16, alpha = 0.2)
```

```{r}
p3
p4
```

- casual counts & registered counts: opposite behaviors - active on weekends vs active on working days. 


## Modeling

- for casual counts: adding temp^2 makes the prediction even more unstable.
- for registered counts: adding temp^2 seems necessary. 

```{r}
# Model 1: starting model, simplest

mod.cas.1 <- lm(casual ~ 
                   workingday + weathersit + atemp + I(atemp^2), data = data2011)
mod.reg.1 <- lm(registered ~ 
                   workingday + weathersit + atemp + I(atemp^2), data = data2011)

# Model 1.5: adding season

mod.cas.1.5 <- lm(casual ~ 
                   workingday + weathersit + atemp + I(atemp^2) + season, data = data2011)
mod.reg.1.5 <- lm(registered ~ 
                   workingday + weathersit + atemp + I(atemp^2) + season, data = data2011)

# Model 2: an intermediate model, slightly more complicated

mod.cas.2 <- lm(casual ~ 
                   weathersit:workingday:atemp + season, data = data2011)
mod.reg.2 <- lm(registered ~ 
                   weathersit:workingday:atemp + I(atemp^2) + weathersit + season, data = data2011)
```

```{r}
# Model 2.5: having season as a interception or an interaction?

mod.cas.2.5 <- lm(casual ~
                   workingday:atemp + weathersit:season, data = data2011)
mod.reg.2.5 <- lm(registered ~
                   workingday:atemp + I(atemp^2) + weathersit + season, data = data2011) 

# adding the holiday variable for cas
mod.cas.2.7 <- lm(casual ~
                   workingday:atemp + weathersit:season + holiday, data = data2011) #current best

# Model 3: having season as an interaction term is better
mod.cas.3 <- lm(casual ~
                  season:workingday:atemp + weathersit + holiday, data = data2011)  #overfitting
mod.reg.3 <- lm(registered ~
                  season:workingday:atemp + season:I(atemp^2) + weathersit, data = data2011)

mod.cas.3.5 <- lm(casual ~
                  season:workingday:atemp + weathersit:season + holiday, data = data2011) #overfitting but leads to best total

# adding the holiday variable for reg
mod.reg.3.7 <- lm(registered ~
                  season:workingday:atemp + season:I(atemp^2) + weathersit + holiday, data = data2011) 


# Model 4: then try a more complex model - but it starts to overfit #alright now no longer overfitting
mod.reg.4 <- lm(registered ~ 
                  season:workingday:atemp + season:workingday:I(atemp^2) + weathersit + holiday, data = data2011) #current best
mod.reg.4.5 <- lm(registered ~ 
                  season:workingday:atemp + season:workingday:I(atemp^2) + weathersit:season + holiday, data = data2011) #overfitting


# current best models
mod.cas.best <- mod.cas.2.7
mod.reg.best <- mod.reg.4

# try removing the y-space outlier
mod.cas.best. <- lm(casual ~
                   season:workingday:atemp + weathersit:season + holiday, 
                   data = filter(data2011, dteday != '2011-08-29'))
mod.reg.best. <- lm(registered ~
                  season:workingday:atemp + season:workingday:I(atemp^2) + weathersit + holiday, 
                  data = filter(data2011, dteday != '2011-08-29'))
```

```{r}
data2011_augmented <- read.csv("data2011_G.csv", header = T)
data2012_augmented <- read.csv("data2012_G.csv", header = T)
```


```{r}
mod.cas.2.7.5 <- lm(casual ~
                   workingday:atemp + weathersit:season + holiday, data = data2011) #current best
mod.cas.3.5.5 <- lm(casual ~
                  season:workingday:atemp + weathersit:season + holiday, data = data2011) #overfitting but leads to best total
mod.reg.4.0.5 <- lm(registered ~ 
                  season:workingday:atemp + season:workingday:I(atemp^2) + weathersit + holiday, data = data2011) #current best
```


```{r}
# guess what, let's try directly modeling the total count! 
mod.tot.1   <- lm(cnt ~ workingday + weathersit + atemp + I(atemp^2), data = data2011)
mod.tot.1.5 <- lm(cnt ~ workingday + weathersit + atemp + I(atemp^2) + holiday + season, data = data2011)
mod.tot.2   <- lm(cnt ~ weathersit:workingday:atemp + I(atemp^2) + weathersit + season, data = data2011)
mod.tot.3   <- lm(cnt ~ season:workingday:atemp + season:I(atemp^2) + weathersit + holiday, data = data2011)
mod.tot.4   <- lm(cnt ~ season:workingday:atemp + season:workingday:I(atemp^2) + weathersit:season, data = data2011)
mod.tot.4.5 <- lm(cnt ~ season:workingday:atemp + season:workingday:I(atemp^2) + weathersit:season + holiday, data = data2011) #cas.2.7 + reg.4
mod.tot.4.7 <- lm(cnt ~ season:workingday:atemp + season:workingday:I(atemp^2) + weathersit:season + holiday, data = data2011) #cas.3.5 + reg.4...oh it's the same...
```


#### Model Evals

```{r message=FALSE, warning=FALSE}
get_mod_eval(mod.cas.1, mod.reg.1,     data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval(mod.cas.1.5, mod.reg.1.5, data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval(mod.cas.2, mod.reg.2,     data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval(mod.cas.2.5, mod.reg.2.5, data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval(mod.cas.2.7, mod.reg.3.7, data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval(mod.cas.3, mod.reg.3,     data2011, filter(data2012, dteday != '2012-10-29')) #casual overfitting
get_mod_eval(mod.cas.3.5, mod.reg.4,   data2011, filter(data2012, dteday != '2012-10-29')) #best total (casual overfitting)
# get_mod_eval(mod.cas.3.5, mod.reg.4.5, data2011, filter(data2012, dteday != '2012-10-29')) #now reg overfitting too

# real current best!!!
get_mod_eval(mod.cas.best,  mod.reg.best,  data2011, filter(data2012, dteday != '2012-10-29')) #best reg + best cas
get_mod_eval(mod.cas.best., mod.reg.best., data2011, filter(data2012, dteday != '2012-10-29')) #outlier removal not really helping
```

```{r}
get_mod_eval_tot(mod.tot.1,data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval_tot(mod.tot.1.5,data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval_tot(mod.tot.2,data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval_tot(mod.tot.3,data2011, filter(data2012, dteday != '2012-10-29'))
get_mod_eval_tot(mod.tot.4,data2011, filter(data2012, dteday != '2012-10-29')) #starts to overfit
get_mod_eval_tot(mod.tot.4.5,data2011, filter(data2012, dteday != '2012-10-29')) #yeaaah separate modeling is better
get_mod_eval_tot(mod.tot.4.7,data2011, filter(data2012, dteday != '2012-10-29')) #yeaaah separate modeling is better
```

```{r}
summary(mod.cas.2.7)
summary(mod.cas.3.5)
summary(mod.reg.4)
```



#### Model 3

- Problem: 
  - underfit in the middle of a week -> might need to incorporate the weekday variable
  - or will need to remove the outlier
  

```{r}
p2 + geom_line(data2011, 
            mapping = aes(dteday, predict(mod.cas.2.7, data2011), color = 'on holidays'), 
            size = 0.5, alpha = 0) + 
  geom_line(data2011, 
            mapping = aes(dteday, predict(mod.cas.3.5, data2011), color = 'fitted'), 
            size = 0.5, alpha = 0.4)

p1 + geom_line(data2011, 
            mapping = aes(dteday, predict(mod.reg.4, data2011), color = 'on holidays'), 
            size = 0.5, alpha = 0.4) + 
  geom_line(data2011, 
            mapping = aes(dteday, predict(mod.reg.4.5, data2011), color = 'fitted'), 
            size = 0.5, alpha = 0.4)

```

```{r}
p4 + geom_line(data2012, 
            mapping = aes(dteday, predict(mod.cas.3, data2012), color = 'fitted'), 
            size = 0.5, alpha = 0) + 
  geom_line(data2012, 
            mapping = aes(dteday, predict(mod.cas.3, data2012)/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.8)

p3 + geom_line(data2012, 
            mapping = aes(dteday, predict(mod.reg.4, data2012)/0.608, color = 'on holidays'), 
            size = 0.5, alpha = 0) + 
  geom_line(data2012, 
            mapping = aes(dteday, predict(mod.reg.3, data2012)/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.8)
```

#### Model 2: previous best

```{r}
p2 + geom_line(data2011, 
            mapping = aes(dteday, predict(mod.cas.2, data2011), color = 'fitted'), 
            size = 0.5, alpha = 0.7)

p1 + geom_line(data2011, 
            mapping = aes(dteday, predict(mod.reg.2, data2011), color = 'fitted'), 
            size = 0.5, alpha = 0.7)

```

```{r}
p4 + geom_line(data2012, 
            mapping = aes(dteday, predict(mod.cas.2, data2012)/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.7)

p3 + geom_line(data2012, 
            mapping = aes(dteday, predict(mod.reg.2, data2012)/0.608, color = 'fitted'), 
            size = 0.5, alpha = 0.7)
```

### Inspecting Data

```{r}
p1 + geom_line(data2011, 
            mapping = aes(dteday, predict(mod.reg.2, data2011), color = 'fitted'), 
            size = 0.5, alpha = 0.7)
```



```{r}
# data2011[data2011$cnt <= 800, ]
# Use filter() from "dplyr (data plyer)" also in tidyverse 
# data2011 %>% filter(cnt <= 800)

# Equivalent code!
filter(data2011, (registered <= 1000) & (dteday > '2011-06-01'))
```

```{r}
# data2012[data2012$cnt <= 1000, ]
```
