---
title: "report4-plots"
author: "Taqi"
date: "11/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Residual Diagostics

### Residual Trends

$$\text{Fig.9: Standardized Residual Plots}$$

```{r fig.width=6, fig.height=2, echo=FALSE}
# Standardized Residuals vs Temperature
StanRes.2 <- rstandard(m.mlr.2)
data.mlr.2 <- data.frame(atemp,StanRes.2)

StanRes.3_1 <- rstandard(m.mlr.3_1)
data.mlr.3_1 <- data.frame(atemp,StanRes.3_1)

StanRes.4_2 <- rstandard(m.mlr.4_2)
data.mlr.4_2 <- data.frame(atemp,StanRes.4_2)

r2 <- ggplot() + 
  geom_point(data=data.mlr.2,   aes(x=atemp, y=StanRes.2,   color = names(cols[2])), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = names(cols[2]), values = cols[2]) +
  labs(x = "Temperature (ºC)", y = "Standarized Residual", title = "Model 2") + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

r3_1 <- ggplot() + 
  geom_point(data=data.mlr.3_1, aes(x=atemp, y=StanRes.3_1, color = names(cols[3])), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = names(cols[3]), values = cols[3]) +
  labs(x = "Temperature (ºC)", y = "Standarized Residual", title = "Model 3_1") +
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

r4_2 <- ggplot() + 
  geom_point(data=data.mlr.4_2, aes(x=atemp, y=StanRes.4_2, color = names(cols[4])), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = names(cols[4]), values = cols[4]) +
  labs(x = "Temperature (ºC)", y = "Standarized Residual", title = "Model 4_2") + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))


plot_grid(r2, r3_1, r4_2, ncol = 3)
```

Most data points spreed evenly above and below the x-axis, which means the predicted values have captured the correct trends in response variation, which is good. Note that for the quadratic model (Model 2), there is still a slight positive bias in the lower-temperature parts and a negative bias in the higher-temperature parts, which also matches our observation in Fig.7&8. This problem is alleviated in the higher-degree models. 

### Residual Normality

#### Standardized Residual versus Fitted Value

$$\text{Fig.10: Standardized Residual Plots}$$

```{r fig.width=6, fig.height=2, echo=FALSE}
# Standardized Residuals vs Fitted
data.mlr.2   <- data.frame(fitted_vals2,   StanRes.2)
data.mlr.3_1 <- data.frame(fitted_vals3_1, StanRes.3_1)
data.mlr.4_2 <- data.frame(fitted_vals4_2, StanRes.4_2)

r2 <- ggplot() + 
  geom_point(data=data.mlr.2,   aes(x=fitted_vals2, y=StanRes.2,   color = names(cols[2])), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = names(cols[2]), values = cols[2]) +
  labs(x = "Predicted Count", y = "Standarized Residual", title = "Model 2") + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

r3_1 <- ggplot() + 
  geom_point(data=data.mlr.3_1, aes(x=fitted_vals3_1, y=StanRes.3_1, color = names(cols[3])), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = names(cols[3]), values = cols[3]) +
  labs(x = "Predicted Count", y = "Standarized Residual", title = "Model 3_1") + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

r4_2 <- ggplot() + 
  geom_point(data=data.mlr.4_2, aes(x=fitted_vals4_2, y=StanRes.4_2, color = names(cols[4])), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = names(cols[4]), values = cols[4]) +
  labs(x = "Predicted Count", y = "Standarized Residual", title = "Model 4_2") + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))


plot_grid(r2, r3_1, r4_2, ncol = 3)
```

When the predicted counts are low, which corresponds to the lower-temperature parts in Fig.10, the residuals have a positive bias, i.e., the predicted values tend to be less than the true values, which matches what we have seen in Fig.7&8. 

The problem that the predicted rental counts go below 0 still needs to be fixed in the later studies. 

#### Normal Q-Q Plots

$$\text{Fig.11: Normal Q-Q Plots}$$

```{r fig.width=6, fig.height=2, echo=FALSE}
# QQ Plot
p2 <- ggplot(data.frame(StanRes.2), aes(sample = StanRes.2, color = names(cols[2]))) + 
  scale_color_manual(name = element_blank(), labels = names(cols[2]), values = cols[2]) + 
  stat_qq() + stat_qq_line() + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

p3_1 <- ggplot(data.frame(StanRes.3_1), aes(sample = StanRes.3_1, color = names(cols[3]))) + 
  scale_color_manual(name = element_blank(), labels = names(cols[3]), values = cols[3]) + 
  stat_qq() + stat_qq_line() + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5)) 

p4_2 <- ggplot(data.frame(StanRes.4_2), aes(sample = StanRes.4_2, color = names(cols[4]))) + 
  scale_color_manual(name = element_blank(), labels = names(cols[4]), values = cols[4]) + 
  stat_qq() + stat_qq_line() + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5)) 

plot_grid(p2, p3_1, p4_2, ncol = 3)
```

All of the residual distributions (Fig.11) match quite well with the standard normal distribution, which is really great, as such degree of matches can rarely be seen in general. 

\newpage

#### Residual Histograms

$$\text{Fig.12: Residual Histograms}$$

```{r fig.width=6, fig.height=2, echo=FALSE}
# Histogram of Quadratic and QMLS

p2 <- ggplot(data.frame(StanRes.2), aes(x = StanRes.2, color = names(cols[2]))) + 
  scale_color_manual(name = element_blank(), labels = names(cols[2]), values = cols[2]) + 
  labs(x = "Standardized Residual", title = "Model 2") + 
  geom_histogram(bins = 100) + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

p3_1 <- ggplot(data.frame(StanRes.3_1), aes(x = StanRes.3_1, color = names(cols[3]))) + 
  scale_color_manual(name = element_blank(), labels = names(cols[3]), values = cols[3]) + 
  labs(x = "Standardized Residual", title = "Model 3_1") + 
  geom_histogram(bins = 100) + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

p4_2 <- ggplot(data.frame(StanRes.4_2), aes(x = StanRes.4_2, color = names(cols[4]))) + 
  scale_color_manual(name = element_blank(), labels = names(cols[4]), values = cols[4]) + 
  labs(x = "Standardized Residual", title = "Model 4_2") + 
  geom_histogram(bins = 100) + 
  theme_grey(base_size = 5.6) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

plot_grid(p2, p3_1, p4_2, ncol = 3)
```

Given the nice match of distributions (Fig.11), the histograms should also match the standard normal distribution quite well, but surprisingly, the cubic model exhibits a slight bi-model issue (Fig.12). This is to be inspected in later analysis. 
